Accurate interactions of cations with neutral and negatively charged membranes via combination of experiments and molecular simulation

### 2018-04-19 Manuscript/manuscript.tex
First pilot content added to the manuscript + figures for both Lipid14/17 and ECC-lipids + some results and references + Makefile for an easy manuscript generation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional macros here.  Please use \newcommand* where
%% possible, and avoid layout-changing macros (which are not used
%% when typesetting).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}

@@

\author{Josef Melcr}
\author{Hector Martinez-Seara}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\author{Ji{\v r}{\' i} Kolafa}
\affiliation{Department of Physical Chemistry, Institute of Chemical Technology, Prague 6, Czech Republic}
\author{Pavel Jungwirth}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\alsoaffiliation{Department of Physics, Tampere University of Technology, P.O. Box 692, FI-33101
Tampere, Finland}

\author{O. H. Samuli Ollila}
\email{samuli.ollila@helsinki.fi}
%\homepage[]{Your web page}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\alsoaffiliation{Institute of Biotechnology, University of Helsinki}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[An \textsf{achemso} demo]
  {Binding of cations to Neutral and Negatively Charged Phospholipid Bilayers 
   made accurate through the use of 
   classical fixed-charge MD models with implicit polarizability }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{MD,IR,NMR,UV}
\keywords{American Chemical Society, \LaTeX}

@@

\begin{abstract}
Binding affinities and stoichiometries of Na$^+$ and Ca$^{2+}$ ions to phospholipid
bilayers are of paramount significance in the properties and functionality of cellular
membranes. Current estimates of binding affinities and
stoichiometries of cations are, however, inconsistent due to limitations in the available experimental and
computational methods. In this work, we improve the description of the binding details of Na$^+$ and Ca$^{2+}$ ions to the
1-Palmitoyl-2-oleoyl-phosphatidylcholine (POPC) bilayer by implicitly including electronic
polarization as a mean field correction, known as the electronic continuum correction (ECC).
This is applied by scaling the partial charges of a selected state-of-the-art POPC lipid model for molecular dynamics simulations.
Our improved ECC-POPC model reproduces not only the
experimentally measured structural parameters for the ion-free membrane, but also the
response of lipid head group to a strongly bound cationic amphiphile, as well as the binding affinities of 
Na$^+$ and Ca$^{2+}$ ions. With our new model we observe on the one side negligible binding of Na$^+$ ions to POPC
bilayer, while on the other side stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens, which is
in agreement with the previous interpretations of the experimental spectroscopic data.
The present model results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
equal probabilities, suggesting more complex binding stoichiometries than those from simple models
used to interpret the NMR data previously.
The results of this work pave the way to quantitative molecular simulations with realistic electrostatic interactions 
of complex biochemical systems at cellular membranes.
\end{abstract}

@@

\section{Introduction}
Interactions of ions with cellular membranes play a key role in many important biological processes~\cite{seelig90,cevc90}. Ions, especially multivalent cations, modify general properties of the membrane which also modulate the embedded transmembrane proteins~\cite{cevc90,tocanne90, pabst07}. For example, Ca$^{2+}$  is crucial in propagation of neural signals. It promotes membrane fusion by bridging lipids in the vesicles carrying neurotransmitters and the neuron synapsis membrane~\cite{papahadjopoulos90,brouwer15}. Ca$^{2+}$ also participates in the T-cell receptor activation. It induces the detachment of the positively charged cytosolic tails of the CD3 protein complex from the negatively charged intracellular membrane making it accessible for the Lck protein~\cite{Shi2013Ca2}. Another example involving sugar-lipid interactions mediated by ions concerns Ca$^{2+}$ modulating the presentation of the sugars present in the PI(4,5)P$_2$ lipid which ultimately modulates the phospholipase C delta 1 pleckstrin homology domain (PLC $\delta$1-H)~\cite{Bilkova2017Calcium}. Interestingly, this effect is not present for Mg$^{2+}$, which illustrates the selectivity of these processes with respect to particular ions. Despite our increasing understanding of the role of ions in cell membrane-related processes the molecular details of the mechanisms behind such processes remain elusive in many cases.

Direct measurements of ion-membrane interactions in complex biological systems are difficult. 
Hence, simplified lipid bilayers are often used as entry-level models to shed light on the role of ions in complex biological membranes~\cite{scherer87,seelig90,cevc90}. 
For this reason, interactions of biologically relevant cations, in particular Na$^+$ and Ca$^{2+}$, with zwitterionic phosphocholine (PC) bilayers have been widely studied experimentally~\cite{akutsu81, altenbach84, seelig90, cevc90, tocanne90, binder02, pabst07, uhrikova08} and using molecular dynamics (MD) simulations~\cite{bockmann03, bockmann04, Berkowitz12, melcrova16, javanainen17}. The details of ion binding are, however, not fully consistent in the literature. Non-invasive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering, and infrared spectroscopies mostly suggest that Na$^+$ ions exhibit negligible binding to PC lipid bilayers. In contrast, Ca$^{2+}$ is observed to specifically bind to a couple of PC molecules via their phosphate groups~\cite{hauser76, hauser78, herbette84, akutsu81, altenbach84, binder02, pabst07, uhrikova08}. Most atomistic resolution MD simulation models predict stronger membrane bindings of the cations than observed in experiments~\cite{catte16}. Namely, simulations report various degrees of Na$^+$ accumulation at the lipid interface~\cite{bockmann03} and for Ca$^{2+}$ strong binding to up to four PC lipids simultaneously, involving not only interactions with phosphates but also with carbonyl oxygens~\cite{bockmann04, melcrova16, javanainen17}.

Recent studies within the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16} made an attempt to resolve these disagreements. A direct comparison of ion binding affinities to PC bilayers between simulations and experiments has been made possible using the electrometer concept~\cite{seelig87}. Namely, the changes in NMR order parameters of the head group upon addition of ions are directly compared to the MD simulations results. 
Analyzing massive amounts of data collected within an open collaboration approach, 
it was concluded that the accuracy of the current state-of-the-art lipid models for MD simulations is not sufficient for a detailed interpretation of the interactions of cations with PC lipid bilayers~\cite{catte16}.

In this work, we improve the description of cation binding to a POPC bilayer within MD simulations 
via an implicit inclusion of electronic polarizability in the polar region of phospholipids. 
To this end, we apply the electronic continuum correction (ECC)~\cite{leontyev11}, 
within which scaling of the atomic partial charges is employed to account for the missing electronic polarizability in standard force field MD simulations. 
Such an approach has been shown previously to improve the behavior of ions in simulations of aqueous salt solutions~\cite{martinek17, Pluharova2014, kohagen14, kohagen16}. Here, we extend the ECC approach to aqueous lipid membranes taking the Lipid14 force field~\cite{dickson14} as a starting point for refining the POPC model. This choice is justified by the fact that the Lipid14 provides the least inaccurate descriptions of cation binding among the existing force fields\cite{catte16}. 
The newly developed ECC-POPC model reproduces the experimentally
measurable structural parameters of an ion-free POPC lipid bilayer with the accuracy comparable to the best state-of-the-art lipid models, while at the same time significantly improving the membrane binding affinities of sodium and calcium cations.

@@

\subsection{Electronic continuum correction for lipid bilayers}\label{section:ecc}
The lack of electronic polarizability in standard MD force fields has been considered a serious issue since the early days of lipid bilayer simulations. In this work, we circumvent the demanding explicit inclusion of electronic polarization effects~\cite{lucas12, chowdhary13} by accounting for the electronic part of polarizability in lipid bilayer simulations implicitly via the electronic continuum correction (ECC)~\cite{leontyev11}. Technically, this is similar to the phenomenological charge-scaling applied in earlier studies of surfactants, lipids or ionic liquids \cite{jonsson86, egberts94, beichel14}. However, the present concept of ECC is physically well justified and has theoretical support~\cite{leontyev09, leontyev10, leontyev11, leontyev14}.

According to ECC, the electronic polarizability can be included into non-polarizable MD in a mean-field way by 
embedding the ions in a homogeneous dielectric continuum with a dielectric constant $\epsilon_{el}$, which is the electronic part of the dielectric constant of the medium~\cite{leontyev11}. Following Coulomb's law,  ECC can be directly incorporated by scaling the charges with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$, yielding
\begin{equation}
Q^{ECC} = f_q \cdot Q
\end{equation}
for the ECC corrected charges. 
Given that the  high frequency dielectric constant of water is $\epsilon _{el} = 1.78$ (\textit{i.e.,} the square of the refraction index), the scaling factor for ions in water is roughly $f_q \approx 0.75$. This scaling factor significantly improves the accuracy of simulations of solvated ions, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014, martinek17}.
It is important to note that the value of the
high frequency dielectric constant 
is around 2 for almost any biologically relevant environment \cite{leontyev11}.
The dielectric discontinuity in a lipid bilayer thus arises only
from the orientational polarization of the molecules, which is accounted for explicitly in standard MD simulations. 
Therefore, the same correction for the electronic polarizability can be 
applied throughout the lipid bilayer/aqueous solution interface.

While using the scaling factor of $f_q = 0.75$ for ions in water is well justified in the ECC theory \cite{leontyev11}, it is not clear whether the same factor should be applied to partial charges used to describe molecules in MD models, e.g., lipids in our case \cite{leontyev14}. Unlike the total charge of an ion, atomic partial charges within molecules are not physical observables. Several computational schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}, with the restrained electrostatic potential method (RESP) being commonly used~\cite{RESP_paper, Singh1984}. Considering that water is often included in RESP calculations or charges are refined to improve certain experimental observables, the electronic polarizability effects of the solvent may to some extent be included in standard force fields~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}. Thus, our application of the scaling factor, $f_q$, to existing partial charges in molecules does not necessarily follow $f_q = \epsilon _{el} ^{-1/2}$. Instead, a consistent scaling factor should lie between the value of 0.75  (\textit{i.e.,} no electronic polarizability included in the original partial charges) and 1 (\textit{i.e.,} electronic polarizability fully included in the original partial charges).

Here we develop a ECC-POPC lipid model that accurately describes binding 
of sodium and calcium ions to the POPC  lipid bilayer. 
The Lipid14~\cite{dickson14} force field 
(available in a Gromacs format from Ref. \citenum{lipid14files}) was used as a starting 
point since 
it provides the most accurate response of the head groups to ions among the available 
lipid models (see Figs.~2~and~5 in Ref.~\citenum{catte16}). Additionally, the Lipid14 model 
provides relatively realistic head group, glycerol backbone, and acyl chain structures~\cite{dickson14,botan15}.
We applied the ECC correction 
to the Lipid14 model of POPC by scaling 
partial charges of the head group, glycerol 
backbone, and carbonyl regions. 
These are the polar parts of phospholipids which can 
contribute to the cation binding.

To reproduce the experimental ion binding affinities,
we scanned possible values of the scaling factor from the interval $f_q~\in~(0.75, 1.0)$.
The ion binding affinity was benchmarked against experiments
using the head group order parameters and the electrometer concept \cite{seelig87,catte16},
as discussed more detail in the next section.
Scaling down the partial charges reduced the ion binding affinity.
We found the most accurate ion binding affinities with a scaling factor of $f_q = 0.8$,
which is only slightly higher than the ECC one ($f_q=0.75$).
Note that common empirical scaling factors for monovalent ions in water are 0.8 or
even closer to unity \cite{benavides17,skinner14,nacleps}.  In contrast, modern force fields
for ionic liquids suggest values of 0.6--0.65, which are lower than $\epsilon^{-1/2}_{el}$\cite{holm14}.
Directly applying
the 0.8 scaling to the partial charges of the head group, the glycerol backbone, and
the carbonyls reduced the area per lipid to 60~\AA$^2$. This area is smaller than in the
original Lipid14 model ($65.6 \pm 0.5$~\AA$^2$)~\cite{dickson14} and in experiments
(64.3~\AA$^2$) \cite{kucerka11}. The decrease of the area per lipid arises from a
reduced hydration of the lipid head group region after scaling of the partial charges, which effectively
reduces the head group polarity. We solve this problem by slightly reducing the effective radii of
the modified head group atoms via lowering the $\sigma$ parameters in the Lennard-Jones potential by a
factor of $f_\sigma = 0.89$. The same approach was successfully adopted for the ECC ions in aqueous
solutions previously \cite{kohagen14, kohagen16, Pluharova2014, martinek17}. After reducing the head group atoms $\sigma$ parameters, the area per molecule is restored to the experimental value (Table~\ref{tab:apls}).

\subsection{Electrometer concept} \label{section:electrometer}
Comparing MD simulation to NMR experiments, we can validate the ion
binding affinity in lipid bilayer simulations using the 'electrometer concept'~ \cite{seelig87, catte16}.
This method is based on the experimental observation that the C-H bond order parameters of $\alpha$ and $\beta$ carbons in a PC lipid head group (Fig.~\ref{simVSexpNOions}) are proportional to the amount of charge bound per lipid~\cite{seelig87}.
The order parameters for all C-H bonds in lipid molecules can be accurately measured using $^2$H NMR or $^{13}$C NMR techniques \cite{ollila16}. From MD simulations the order parameters can be calculated using the relation
\begin{equation}\label{OP}
S_{\rm CH}=\frac{1}{2}\langle 3\cos^2\theta -1 \rangle,
\end{equation}
where $\theta$ is the angle between the C-H bond and the membrane normal. Angular brackets point to the average over all sampled configurations.

The relation between the amount of the bound charge per lipid, $X^\pm$, and the head group order parameter change, $\Delta S_{\rm{CH}}^{i}$, is empirically quantified as~\cite{seelig87,ferreira16}
\begin{equation}\label{OPchangeEQ}
\Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) \approx m_i \frac{4 }{3\chi}X^\pm,
\end{equation}
where $i$ refers to either the $\alpha$ or $\beta$ carbons, $S_{\rm{CH}}^{i}(0)$ denotes the order parameter in the absence of bound charge, $\chi$ is the quadrupole coupling constant ($\chi \approx$\,167\,kHz), and $m_i$ is an empirical constant depending on the valency and location of the bound charge.

The measured change of the order parameter depends on the head group response to the bound charge and on the amount of the bound charge (\textit{i.e.,} $m_i$ and $X^\pm$ in Eq.~\ref{OPchangeEQ}, respectively).  The empirical factor $m_i$ has to be adequately quantified before the electrometer concept can be used to analyze the binding affinities. This calibration has been done experimentally for a wide range of systems~\cite{seelig87, beschiasvili91}. To calibrate the response of the head group order parameters to the bound charge in simulations, we use experimental data for a strong cationic surfactant dihexadecyldimethylammonium bromide  (DHAB) mixed with a POPC bilayer~\cite{scherer89}. DHAB\\[0.5cm]
%\vspace{0.5cm}
\chemfig{
  -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:150]H_3C)(-[:210]H_3C)(-[:330]{(}CH_2{)}_{15}-CH_3)(-[:30]{(}CH_2{)}_{15}-CH_3)
}  
\vspace{0.5cm} \\
is a cationic surfactant having two acyl chains and bearing a unit charge at the hydrophilic end. Due to its structure it is expected to locate in the bilayer similarly to the phospholipids and its molar ratio then gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems~\cite{scherer89}.

\subsection{Salt concentrations and binding affinities}
When measuring head group order parameters, the NMR experiments report the employed salt concentrations in two different ways.
Some studies provide the salt concentrations in water before solvating the lipids~\cite{akutsu81}, $C_{ion}'$,
while others use atomic absorption spectroscopy and report the salt concentration in
the supernatant after the solvation of lipids~\cite{altenbach84}, $C_{ion}$. In this work,
we stick to the latter definition by estimating the salt concentration in the aqueous
bulk region using the farthest point from the lipid bilayer in the aqueous phase.
Note that the former definition was used by some of us previously.~\cite{catte16}.
Although these two definitions provide somewhat different values when applied
to Ca$^{2+}$ concentrations, their particular choice does not affect the conclusions of this work in any significant way.

To quantify the ion binding affinities to a lipid bilayer, we calculate the relative surface excess of ions with respect to water, $\Gamma_{ion}^{water}$~\cite{chattorajBOOK}. Such a quantity compares the adsorption of ions to the adsorption of water molecules at the interface without the necessity of defining a Gibbs dividing surface between the membrane interior and the water bulk region. In our simulations, we only assume that the interface is located between the ion-free hydrophobic interior of the lipid bilayer and the aqueous region far from the membrane.  Such a setup and the above definition of bulk ion concentrations provides a simple relation for the relative surface excess $\Gamma_{ion}^{water}$ for simulations of lipid bilayers,
 \begin{equation}\label{surfexcess}
  \Gamma_{ion}^{water}=\frac{1}{2A_b} \left ( n_{ion} - n_{water} \frac{C_{ion}}{C_{water}} \right ).
\end{equation}
Here, $n_{water}$ and $n_{ion}$ are the total numbers of water molecules and ions in the system,
$C_{water}$ and $C_{ion}$ are their respective bulk concentrations in the aqueous phase
(estimated from the solution density of farthest point between periodic bilayers),
and $A_b$ is the area of the unit cell in the membrane plane.
The total area of the interface is then twice the area of the membrane, i.e., $2A_b$,
because the bilayer has an interface at each of the two leaflets.

\subsection{Validation of lipid bilayer structure against NMR and scattering experiments}
The structures of lipid bilayers in simulations without ions were validated against NMR by calculating the order parameters for the C-H bonds and against \mbox{X-ray} scattering experiments by evaluating the scattering form factors. NMR order parameters validate the structures sampled by the individual lipid molecules with atomic resolution. The simulated order parameters were calculated for all C-H bonds in lipid molecules from Eq. \ref{OP}. Scattering form factors validate the dimensions of the lipid bilayer (i.e., the bilayer thickness and area per molecule). Form factors were calculated using a relation

@@

\subsubsection{Simulations of POPC bilayers with aqueous ions}
Simulations of a POPC bilayer in pure water or at varying salt concentrations consisted of 128 POPC molecules
and approximately 50~water molecules per lipid in an orthorhombic simulation box with periodic boundary conditions.
The SPC/E~\cite{Berendsen1987} water model was used in all ECC-POPC model simulations reported
in the main text. This model was also used in the previous parametrization
of ECC-ions \cite{martinek17, kohagen16, Pluharova2014} because its lowered dielectric constant is consistent with
the ECC concept~\cite{leontyev11, leontyev14}.
The robustness of our approach with other water models (OPC \cite{Izadi14},
OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3P-FB and TIP4P-FB~\cite{Wang2014}, and TIP4P/2005 \cite{Abascal2005})
was verified and is presented in the Supporting Information (SI).
Sodium, calcium, and chloride ions were modeled as ECC-ions
with parameters from Refs. \citenum{martinek17, Pluharova2014}. Scaled charges and Lennard-Jones radii for atoms
forming the POPC lipid were derived in this work starting from the Lipid14 force field~\cite{dickson14}.
For comparison, simulations with the Lipid14 model~\cite{dickson14} and
ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006} or ECC-ions \cite{martinek17, kohagen16, Pluharova2014}
were also performed. The TIP3P water model~\cite{jorgensen83} was used in all simulations with the original Lipid14 model.
Simulation data for the Lipid14 model with \AA{}qvist ions~\cite{aqvist90} and the TIP3P~\cite{jorgensen83} water model were taken directly from
Refs.~\citenum{catte16,lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC,lipid14POPC1000mMNaClfiles}.

MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4). The simulation parameters used
in this work are summarized in Table~S1 in the SI. 
Compatibility with the openMM simulation package \cite{openmm7} was also
tested and is presented in the SI. 
For deposited trajectories and parameter files see 
Refs.~\citenum{lipid14_POPC_Dang_and_ECC-ions_files, ECC-POPC_nacl_cacl2_files, ECC-POPC_nacl_cacl2_wat-models_files}.

\subsubsection{Simulations of POPC bilayers with cationic surfactants}
The topology of the dihexadecyldimethylammonium cation was created with the automated topology builder \cite{malde11}. The AmberTools program~\cite{amber} was used to generate the Amber-type force field parameters. These parameters were then converted to the Gromacs format with the acpype tool \cite{acpype}. The partial charges were manually modified in the non-scaled version to approximately match similar segments in Lipid14 \cite{dickson14}. For simulation with ECC-POPC, the partial charges of the head group of the cation were scaled by a factor of $f_q=0.8$, and the $\sigma$ of the scale atoms reduced by a factor of 0.89 similarly as for POPC-ECC.

The cationic surfactants were randomly mixed among the phospholipids to form bilayer structures with mole fractions of 10\%, 20\%, 30\%, 42\%, or 50\% of surfactant in the POPC bilayer. All these systems contained 50 POPC molecules per leaflet, 6340 water molecules, and 6, 14, 21, 35, or 50 surfactants per leaflet.  Chloride counter ions were used in all the simulations, while bromide was the counter ion in the experiment~\cite{scherer89}. For POPC, either Lipid14 with TIP3P~\cite{jorgensen83} or ECC-POPC with SPC/E~\cite{Berendsen1987} parameters  were used. The first 20~ns of the total simulation time of 200~ns was considered as an equilibration period and was thus omitted from the analysis. We checked that a sufficient lipid  exchange occurred during the simulations.
For deposited trajectories and parameter files of scaled and unscaled simulations see
Refs.~\citenum{ECC-POPC_cationic-surf_files} and~\citenum{POPClipid14T313K,POPClipid1410perCATsurfT313K,POPClipid1420perCATsurfT313K,POPClipid1430perCATsurfT313K,POPClipid1442perCATsurfT313K,POPClipid1450perCATsurfT313K}, respectively.

@@

\subsection{Structural parameters of a pure ECC-POPC model membrane: Agreement with experiments}

\begin{figure}[tb!]
  \centering
  \includegraphics[width=8.2cm]{../Fig/ipython_nb/Order-parameters_form-factors_exp-L14-ECCL17_q80_sig89_POPC-struct.pdf}
  \caption{\label{simVSexpNOions}
    Top: X-ray scattering form factors from simulations with the Lipid14 \cite{dickson14} and
    the ECC-POPC models compared with experiments~\cite{kucerka11} at 303~K. \\
    Middle: Order parameters of POPC head group, glycerol backbone and acyl chains 
    from simulations with the Lipid14 \cite{dickson14} and the ECC-POPC models
    compared with experiments \cite{ferreira13} at 300~K.
    The size of the markers for the head group order parameters correspond to
    the error estimate $\pm 0.02$ for experiments \cite{botan15,ollila16},
    while the error estimate for simulations is $\pm 0.005$
    (Bayesian estimate of 95\% confidence interval \cite{scipy}).
    The size of the points for acyl chains are decreased by a factor of 3 to improve the clarity of the plot.
    Open/closed symbols are used for palmitoyl/oleoyl chains of POPC. \\
    Bottom: The chemical structure of POPC and the labeling of the carbon segments.
  } 
\end{figure}

\begin{table}[tb!]
  \caption{Values of the area per lipid (APL) of POPC bilayers without ions. \label{tab:apls}
  }
  \begin{tabular}{l|c c}
    model          & APL (\AA$^2$)   & Temperature [K] \\
    \hline
    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
    \hline
    ECC-POPC                & 63.2$\pm$ 0.6  &  300       \\
    \hline
    experiment \cite{kucerka11} & 64.3  &  303    \\
    \hline
  \end{tabular}
\end{table}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%% Place any additional macros here.  Please use \newcommand* where 
%% possible, and avoid layout-changing macros (which are not used 
%% when typesetting). 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\newcommand*\mycommand[1]{\texttt{\emph{#1}}} 
\newlength{\figwidth}
\setlength{\figwidth}{8 cm}

@@

\author{Josef Melcr} 
\email{melcr@uochb.cas.cz} 
%%\homepage[]{https://jmelcr.github.io/}
\affiliation{Institute of Organic Chemistry and Biochemistry, 
Academy of Sciences of the Czech Republic,  
Prague 6, Czech Republic} 
\author{Pavel Jungwirth} 
%%\homepage[]{http://jungwirth.uochb.cas.cz/}
\affiliation{Institute of Organic Chemistry and Biochemistry, 
Academy of Sciences of the Czech Republic,  
Prague 6, Czech Republic} 
\alsoaffiliation{Department of Physics, Tampere University of Technology, P.O. Box 692, FI-33101 
Tampere, Finland}

\author{O. H. Samuli Ollila} 
\email{samuli.ollila@helsinki.fi} 
%%\homepage[]{Your web page} 
\affiliation{Institute of Organic Chemistry and Biochemistry, 
Academy of Sciences of the Czech Republic,  
Prague 6, Czech Republic} 
\alsoaffiliation{Institute of Biotechnology, University of Helsinki}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%% The document title should be given as usual. Some journals require 
%% a running title from the author: this should be supplied as an 
%% optional argument to \title. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\title[] 
  {Accurate interactions of cations with 
   neutral and negatively charged membranes 
   via combination of experiments and molecular simulation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%% Some journals require a list of abbreviations or keywords to be 
%% supplied. These should be set up here, and will be printed after 
%% the title and author information, if needed. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\abbreviations{IR,NMR,UV,MD,ECC,PC,PS,POPS,POPS} 
\keywords{MD simulation, molecular modeling, 
          polarizability, polarization,
          phospholipids, phosphatidylserine}

@@

\begin{abstract} 
Binding affinities and stoichiometries of sodium and calcium ions to phospholipid bilayers are of paramount significance in the properties and functionality of cellular membranes. Current estimates of binding affinities and stoichiometries of cations are, however, controversial due to limitations in the available experimental and theoretical methods. Experimentally one can assess these parameters by titrating several membrane properties as a function of the ion concentration. However, the interpretation of the experiments relies on theoretical models, as direct molecular information is not available. 
Classical molecular dynamics (MD) simulations provide details of the ion binding process with atomistic resolution, therefore offering all the necessary information to interpret experimental data without the need to resort to over-simplified models. However, the accuracy of the currently available lipid models when interacting with ions is not sufficient for such interpretation \cite{catte16}. Recently, we have improved the binding details of sodium and calcium ions to the 1-Palmitoyl-2-oleoyl-phosphatidylcholine (POPC) bilayer by implicitly including electronic polarization as a mean field correction, known as the electronic continuum correction (ECC). Our improved ECC-POPC model reproduces not only the experimentally measured structural parameters for the ion-free membrane, but also improves significantly the response of lipid head group to a bound positive charge, and the binding affinities of sodium and calcium ions. \cite{melcr18}
We are now extending our work to other lipid membranes composed of larger varieties of different biologically relevant lipids. In particular, we focus our attention on membranes with phosphatidylethanolamine (PE) and negatively charged phosphatidylserine (PS). We compare available NMR and x-ray experiments with MD simulations to obtain atomistic-resolution picture of the interactions between cations and phospholipid membranes. Our preliminary results indicate that electronic polarization is a non-negligable interaction, which dramatically improves the description of cation binding to the membranes. 
\end{abstract}

@@

\section{Introduction}

@@

\subsection{Electronic continuum correction for lipid bilayers}\label{section:ecc} 
This section may not be required, we may just cite ECC-POPS 
and put a short paragraph describing what we did on POPS.

Compared to ECC-POPC model \cite{melcr18}, 
which used used a scaling factor $f_q = 0.8$ for all charges but lipid tails,
this model, ECC-POPS, has to adopt a scaling factor $f_q = 0.75$ 
as it has non-zero total charge in line with ECC theory. \cite{leontyev09}
Hence, electronic continuum correction of POPS 
has the same characteristics as for cations and anions, 
denoted as ECC-ions \cite{martinek17, Pluharova2014, kohagen14, kohagen16}.

\subsection{Electrometer concept} \label{section:electrometer} 
This section may not be required, 
we may just cite ECC-POPS and NMRlipids projects papers (and ongoing works)

\subsection{Validation of lipid bilayer structure against NMR and scattering experiments} 
\emph{This section probably has to stay. Following text to be modified. Taken from Ref.~\citenum{melcr18}}

The structures of lipid bilayers in simulations without ions were validated against NMR by calculating the order parameters for the C-H bonds and against \mbox{X-ray} scattering experiments by evaluating the scattering form factors. 
NMR order parameters validate the structures sampled by the individual lipid molecules with atomic resolution. 
The simulated order parameters were calculated for all C-H bonds in lipid molecules from Eq. \ref{OP}.
Scattering form factors validate the dimensions of the lipid bilayer (i.e., the bilayer thickness and area per molecule). Form factors were calculated using a relation

@@

General simulation setup.

\subsubsection{Simulations of POPS:POPS bilayers with aqueous ions}

Any specifics for simulations with cations.

\subsubsection{Simulations of POPS:POPS bilayers with anionic/cationic surfactants/peptides}

Any specifics for simulations with anionic/cationinc surfactants/peptides. 
e.g. where the parameters for them came from.

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
% ++++++++  RESULTS AND DISCUSSION section +++++++++++++++++++++++++++++++++++++++++++
% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

@@

\subsection{Structural parameters of a pure ECC-POPS model membrane: Agreement with experiments}

\begin{figure}[tb!] 
  \centering 
  \includegraphics[width=\figwidth]{../Fig/order_parameters_actual_pure-POPS.pdf} \\
  \includegraphics[width=0.7\figwidth]{../Fig/form-f_l17-ecc-pops-exp_compar.pdf} 
  \includegraphics[width=0.55 \figwidth]{../Fig/lipids_chemfig.pdf} 
  \caption{\label{simVSexpNOions} 
    X-ray scattering form factors from simulations with the Lipid17 \cite{lipid17-future} and 
    the ECC-POPS models compared with experiments~\cite{SDP-CHARMM36_comparison_paper_Samuli-knows} at 298~K. 
    The disagreements between ECC-POPS model and experiment 
    may come from an intrinsincally incorrect structure of the head group (and, hence, electron density in the region), 
    however, lipid packing and overall membrane properties 
    may be still caputured well enough -- after considering areas per lipid reported in Table~\ref{tab:apls}.  \\
    Order parameters of POPS head group, glycerol backbone and acyl chains  
    from simulations with the Lipid17 \cite{lipid17-future} and the ECC-POPS models 
    compared with \emph{experiments at 300~K (check and change)}. 
    Open/closed symbols are used for palmitoyl/oleoyl chains of POPS. \\
    The chemical structure of POPS and the labeling of the carbon segments. 
  }  
\end{figure}

\begin{table}[tb!] 
  \caption{Values of the area per lipid (APL) of POPS bilayers with only \ce{Na^+} counterions. \label{tab:apls} 
  } 
  \begin{tabular}{l|c c} 
    model          & APL (\AA$^2$)   & Temperature [K] \\ 
    \hline 
    Lipid14                   & 53.5$\pm$ 0.8  &  298 \\ 
    \hline 
    ECC-POPS                & 60.3$\pm$ 0.6  &  298       \\ 
    \hline 
    experiment \cite{SDP-CHARMM36_comparison_paper_Samuli-knows} & 62.3  &  298    \\ 
    \hline 
  \end{tabular} 
\end{table}

\subsection{Electrometer concept in negatively charged membranes with various cations}

Preliminary results are plotted in the figures in this section. 
Beware! -- the y-axis scales are variable!
Colours:
\begin{itemize}
  \item Red -- ECC-lipids, ECC-ions
  \item Blue -- Lipid14/17, Dang ions
\end{itemize}

Changes of the lipid bilayer head group order parameters extracted from simulations and 
experiments \cite{roux90} are shown in Figs.~\ref{fig:delta_ordPar_KCl}, \ref{fig:delta_ordPar_NaCl} 
and~\ref{fig:delta_ordPar_CaCl} as functions of KCl, NaCl or CaCl$_2$ concentrations,
and in Figs.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix} as a function of PC and PS content.
